---
title: "Modelling"
output: html_document
date: "2026-04-10"
---

```{r}
# Libraries
library(rminer)
library(xgboost)
library(ggplot2)
library(tidyverse)
library(dplyr)
```

```{r}
# Get Data
baltimore_uni <- readRDS("data_clean/baltimore_uni.rds")

lancaster_uni <- readRDS("data_clean/lancaster_uni.rds")

philadelphia_uni <- readRDS("data_clean/philadelphia_uni.rds")

richmond_uni <- readRDS("data_clean/richmond_uni.rds")
```

```{r}
# Crate lists of dataframes for each type of problem (Multivariate and Univariate)
datasets_uni <- list(baltimore_uni = baltimore_uni,
                     lancaster_uni = lancaster_uni,
                     philadelphia_uni = philadelphia_uni,
                     richmond_uni = richmond_uni
                     )
```


```{r}
run.xgboost <- function(df, horizon = 7, iter = 20, city_name, params) {
  cat("---- Start Training for XGBoost: ", city_name, "----\n")
  importance_list <- list()
  models_list <- list()
  
  N <- nrow(df)
  window <- N - (iter * horizon) - horizon

  predictions <- list()

  all_nmae <- matrix(NA, nrow = iter, ncol = horizon)
  all_rmse <- matrix(NA, nrow = iter, ncol = horizon)

  for (i in 1:iter) {
    cat("\nIter:", i, "\n")
    index <- holdout(
      1:N,
      ratio = horizon,
      iter = i,
      mode = "incremental",
      increment = horizon,
      window = window
    )

    train_idx <- index$tr

    train <- df[train_idx, ]

    X_train <- as.matrix(train[, !(names(train) %in% c("Date", "Num_Customers")), drop = FALSE])
    y_train <- train$Num_Customers

    d_train <- xgboost::xgb.DMatrix(data = X_train, label = y_train)

    xgb_model <- xgboost::xgb.train(
      params = params,
      data = d_train,
      nrounds = 1000
    )

    models_list[[i]] <- xgb_model

    importance <- xgboost::xgb.importance(model = xgb_model)
    importance$iter <- i
    importance_list[[i]] <- importance

    for (h in 1:horizon) {
      cat("\nHorizon:", h, "\n")
      test_idx  <- index$ts[h]
      test  <- df[test_idx, ]

      x_test  <- as.matrix(test[, !(names(test) %in% c("Date", "Num_Customers")), drop = FALSE])
      y_test  <- test$Num_Customers

      d_test <- xgboost::xgb.DMatrix(data = x_test, label = y_test)
      pred <- as.numeric(predict(xgb_model, d_test))

      predictions[[length(predictions) + 1]] <- data.frame(
        city = city_name,
        iter = i,
        horizon = h,
        date = test$Date,
        actual = y_test,
        predicted = pred,
        error = y_test - pred
      )

      mae  <- abs(y_test - pred)
      nmae <- mae / mean(y_train)
      rmse <- (y_test - pred)^2

      all_nmae[i, h] <- nmae
      all_rmse[i, h] <- rmse
    }
  }

  predictions_df <- do.call(rbind, predictions)

  nmae_per_horizon <- colMeans(all_nmae)
  rmse_per_horizon <- sqrt(colMeans(all_rmse))

  importance_df <- do.call(rbind, Filter(Negate(is.null), importance_list))

  importance_mean <- aggregate(Gain ~ Feature, data = importance_df, mean)
  names(importance_mean)[2] <- "Gain_mean"

  importance_sd <- aggregate(Gain ~ Feature, data = importance_df, sd)
  names(importance_sd)[2] <- "Gain_sd"
  
  importance_freq <- aggregate(Gain ~ Feature, data = importance_df, length)
  names(importance_freq)[2] <- "Frequency"

  importance_summary <- Reduce(function(x, y) merge(x, y, by = "Feature"),
                               list(importance_mean, importance_sd, importance_freq))

  importance_summary <- importance_summary[order(-importance_summary$Gain_mean), ]

  cat("\n---- XGBoost Completed:", city_name, "----\n")

  return(list(
    metrics = list(
      NMAE = nmae_per_horizon,
      RMSE = rmse_per_horizon
    ),
    predictions = predictions_df,
    importance = importance_df,
    importance_summary = importance_summary,
    models = models_list
  ))
}
```


```{r}
# Start with the Univariate Problem
out <- list()

params <- list(
    objective = "reg:squarederror",
    max_depth = 6,
    eta = 0.05,
    subsample = 0.8,
    colsample_bytree = 0.7,
    min_child_weight = 5,
    gamma = 0.1,
    lambda = 1,
    alpha = 0.5,
    early_stopping_rounds = 50
  )

for (city_name in names(datasets_uni)) {
  df <- datasets_uni[[city_name]]

  out[[city_name]] <- run.xgboost(
    df = df,
    horizon = 7,
    iter = 20,
    city_name = city_name,
    params = params
  )
}

all_predictions <- do.call(rbind, lapply(out, function(x) x$predictions))
```

```{r}
compute_weighted_score <- function(df, w_nmae = 0.4, w_rmse = 0.6) {
  df %>%
    group_by(horizon) %>%
    mutate(
      NMAE_norm = ifelse(max(NMAE) == min(NMAE), 0,
                         (NMAE - min(NMAE)) / (max(NMAE) - min(NMAE))),

      RMSE_norm = ifelse(max(RMSE) == min(RMSE), 0,
                         (RMSE - min(RMSE)) / (max(RMSE) - min(RMSE)))
    ) %>%
    ungroup() %>%
    mutate(
      score = w_nmae * NMAE_norm + w_rmse * RMSE_norm
    )
}
```


```{r, fig.width=18, fig.height=8}
# Actual vs Predicted
for (city_name in unique(all_predictions$city)) {
  df_city <- subset(all_predictions, city == city_name)
  
  df_long <- df_city %>%
    pivot_longer(cols = c(actual, predicted),
                 names_to = "type",
                 values_to = "Value")

  print(ggplot(df_long, aes(x = date, y = Value, color = type)) +
    geom_line() +
    facet_wrap(~ horizon, scales = "free_y") +
    ggtitle(paste("City:", city_name)) +
    theme_minimal()
  )
}
```

The model shows good performance in the short term (horizon 1), being able to follow the overall trend of the series. However, as the forecast horizon increases, the prediction quality gradually decreases.

It is observed that the model captures the existence of peaks, but often fails in their timing and magnitude, predicting peaks when none occur and missing them when they actually happen.

This behavior indicates a strong reliance on past patterns (lags), without the ability to anticipate specific events.

The model learns the base structure of the series but fails to explain abrupt variations, indicating the need for improved feature engineering.

```{r, fig.width=18, fig.height=8}
# Error Over Time
for (city_name in names(out)) {
  df_city <- out[[city_name]]$predictions
  
  print(ggplot(df_city, aes(x = date, y = error)) +
    geom_line(color = "red") +
    facet_wrap(~ horizon, scales = "free_y") +
    ggtitle("City:", city_name) +
    theme_minimal()
  )
}
```

```{r, fig.width=18, fig.height=8}
# NMAE per Horizon
for (city_name in names(out)) {
  df_metrics <- data.frame(
    horizon = 1:7,
    NMAE = out[[city_name]]$metrics$NMAE
  )

  print(ggplot(df_metrics, aes(x = horizon, y = NMAE)) +
          geom_line(color = "red") + geom_point() +
          theme_minimal() +
          ggtitle("City:", city_name)
  )
}
```

```{r, fig.width=18, fig.height=8}
# RMSE per Horizon
for (city_name in names(out)) {
  df_metrics <- data.frame(
    horizon = 1:7,
    RMSE = out[[city_name]]$metrics$RMSE
  )

  print(ggplot(df_metrics, aes(x = horizon, y = RMSE)) +
          geom_line(color = "red") + geom_point() +
          theme_minimal() +
          ggtitle("City:", city_name)
  )
}
```

```{r, fig.width=18, fig.height=8}
# Feature Importance
for (city_name in names(out)) {
  top_features <- head(out[[city_name]]$importance_summary, 10)
  
  print(ggplot(top_features, aes(x = reorder(Feature, Gain_mean), y = Gain_mean)) +
    geom_col() +
    coord_flip() +
    ggtitle("City:", city_name) +
    theme_minimal()
  )
}
```

The model shows strong performance for short-to-medium horizons (h=3–5), but struggles significantly for longer horizons, particularly at h=7, where errors increase sharply.

Feature importance analysis reveals that predictions are primarily driven by short-term lags (lag1 and lag7), indicating that the model relies heavily on recent observations and weekly seasonality.

This explains the model's inability to capture sudden spikes, as it lacks contextual or exogenous information that could signal extreme events.

As a result, the model performs well under normal conditions but systematically underestimates peak values, leading to large prediction errors.

