---
title: "LSTM"
output: html_document
date: "2026-04-25"
---

```{r}
# Libraries
library(rminer)
library(xgboost)
library(ggplot2)
library(tidyverse)
library(dplyr)
```

```{r}
# Get Date
baltimore_multi <- readRDS("data_clean/baltimore_multi.rds")

lancaster_multi <- readRDS("data_clean/lancaster_multi.rds")

philadelphia_multi <- readRDS("data_clean/philadelphia_multi.rds")

richmond_multi <- readRDS("data_clean/richmond_multi.rds")
```

```{r}
# Crate lists of dataframes for each type of problem (Multivariate)
datasets_uni <- list(baltimore_multi = baltimore_multi,
                     lancaster_multi = lancaster_multi,
                     philadelphia_multi = philadelphia_multi,
                     richmond_multi = richmond_multi
                     )

```

```{r}
run.xgboost <- function(df, horizon = 7, iter = 20, city_name, params) {
  cat("---- Start Training for XGBoost: ", city_name, "----\n")
  importance_list <- list()
  models_list <- list()
  
  N <- nrow(df)
  window <- N - (iter * horizon) - horizon

  predictions <- list()

  all_nmae <- matrix(NA, nrow = iter, ncol = horizon)
  all_rmse <- matrix(NA, nrow = iter, ncol = horizon)

  for (i in 1:iter) {
    cat("\nIter:", i, "\n")
    index <- holdout(
      1:N,
      ratio = horizon,
      iter = i,
      mode = "incremental",
      increment = horizon,
      window = window
    )

    train_idx <- index$tr

    train <- df[train_idx, ]

    X_train <- as.matrix(train[, !(names(train) %in% c("Date", "Num_Customers")), drop = FALSE])
    y_train <- train$Num_Customers

    d_train <- xgboost::xgb.DMatrix(data = X_train, label = y_train)

    xgb_model <- xgboost::xgb.train(
      params = params,
      data = d_train,
      nrounds = 1000
    )

    models_list[[i]] <- xgb_model

    importance <- xgboost::xgb.importance(model = xgb_model)
    importance$iter <- i
    importance_list[[i]] <- importance

    for (h in 1:horizon) {
      cat("\nHorizon:", h, "\n")
      test_idx  <- index$ts[h]
      test  <- df[test_idx, ]

      x_test  <- as.matrix(test[, !(names(test) %in% c("Date", "Num_Customers")), drop = FALSE])
      y_test  <- test$Num_Customers

      d_test <- xgboost::xgb.DMatrix(data = x_test, label = y_test)
      pred <- as.numeric(predict(xgb_model, d_test))

      predictions[[length(predictions) + 1]] <- data.frame(
        city = city_name,
        iter = i,
        horizon = h,
        date = test$Date,
        actual = y_test,
        predicted = pred,
        error = y_test - pred
      )

      mae  <- abs(y_test - pred)
      nmae <- mae / mean(y_train)
      rmse <- (y_test - pred)^2

      all_nmae[i, h] <- nmae
      all_rmse[i, h] <- rmse
    }
  }

  predictions_df <- do.call(rbind, predictions)

  nmae_per_horizon <- colMeans(all_nmae)
  rmse_per_horizon <- sqrt(colMeans(all_rmse))

  importance_df <- do.call(rbind, Filter(Negate(is.null), importance_list))

  importance_mean <- aggregate(Gain ~ Feature, data = importance_df, mean)
  names(importance_mean)[2] <- "Gain_mean"

  importance_sd <- aggregate(Gain ~ Feature, data = importance_df, sd)
  names(importance_sd)[2] <- "Gain_sd"
  
  importance_freq <- aggregate(Gain ~ Feature, data = importance_df, length)
  names(importance_freq)[2] <- "Frequency"

  importance_summary <- Reduce(function(x, y) merge(x, y, by = "Feature"),
                               list(importance_mean, importance_sd, importance_freq))

  importance_summary <- importance_summary[order(-importance_summary$Gain_mean), ]

  cat("\n---- XGBoost Completed:", city_name, "----\n")

  return(list(
    metrics = list(
      NMAE = nmae_per_horizon,
      RMSE = rmse_per_horizon
    ),
    predictions = predictions_df,
    importance = importance_df,
    importance_summary = importance_summary,
    models = models_list
  ))
}
```

```{r}
# Start with the Univariate Problem
out <- list()

params <- list(
    objective = "reg:squarederror",
    max_depth = 6,
    eta = 0.05,
    subsample = 0.8,
    colsample_bytree = 0.7,
    min_child_weight = 5,
    gamma = 0.1,
    lambda = 1,
    alpha = 0.5,
    early_stopping_rounds = 50
  )

for (city_name in names(datasets_uni)) {
  df <- datasets_uni[[city_name]]

  out[[city_name]] <- run.xgboost(
    df = df,
    horizon = 7,
    iter = 20,
    city_name = city_name,
    params = params
  )
}

all_predictions <- do.call(rbind, lapply(out, function(x) x$predictions))
```

```{r}
compute_weighted_score <- function(df, w_nmae = 0.4, w_rmse = 0.6) {
  df %>%
    group_by(horizon) %>%
    mutate(
      NMAE_norm = ifelse(max(NMAE) == min(NMAE), 0,
                         (NMAE - min(NMAE)) / (max(NMAE) - min(NMAE))),

      RMSE_norm = ifelse(max(RMSE) == min(RMSE), 0,
                         (RMSE - min(RMSE)) / (max(RMSE) - min(RMSE)))
    ) %>%
    ungroup() %>%
    mutate(
      score = w_nmae * NMAE_norm + w_rmse * RMSE_norm
    )
}
```

```{r, fig.width=18, fig.height=8}
# Actual vs Predicted
for (city_name in unique(all_predictions$city)) {
  df_city <- subset(all_predictions, city == city_name)
  
  df_long <- df_city %>%
    pivot_longer(cols = c(actual, predicted),
                 names_to = "type",
                 values_to = "Value")

  print(ggplot(df_long, aes(x = date, y = Value, color = type)) +
    geom_line() +
    facet_wrap(~ horizon, scales = "free_y") +
    ggtitle(paste("City:", city_name)) +
    theme_minimal()
  )
}
```

```{r, fig.width=18, fig.height=8}
# Error Over Time
for (city_name in names(out)) {
  df_city <- out[[city_name]]$predictions
  
  print(ggplot(df_city, aes(x = date, y = error)) +
    geom_line(color = "red") +
    facet_wrap(~ horizon, scales = "free_y") +
    ggtitle("City:", city_name) +
    theme_minimal()
  )
}
```

```{r, fig.width=18, fig.height=8}
# NMAE per Horizon
for (city_name in names(out)) {
  df_metrics <- data.frame(
    horizon = 1:7,
    NMAE = out[[city_name]]$metrics$NMAE
  )

  print(ggplot(df_metrics, aes(x = horizon, y = NMAE)) +
          geom_line(color = "red") + geom_point() +
          theme_minimal() +
          ggtitle("City:", city_name)
  )
}
```

The figure shows the NMAE evolution across different forecast horizons for each city using a multivariate approach.

Overall, the model performs best at intermediate horizons (h=3 to h=5), where the lowest errors are observed. This suggests that the model captures short-term temporal dynamics effectively while still maintaining stability in the medium term.

At shorter horizons (h=1–2), the error is slightly higher, indicating sensitivity to recent fluctuations. At longer horizons (especially h=7), the error increases, which is expected due to the accumulation of uncertainty over time.

Despite some variability across cities, the model remains consistent, with most NMAE values within a good performance range (~0.07–0.18), indicating solid predictive capability.

```{r, fig.width=18, fig.height=8}
# RMSE per Horizon
for (city_name in names(out)) {
  df_metrics <- data.frame(
    horizon = 1:7,
    RMSE = out[[city_name]]$metrics$RMSE
  )

  print(ggplot(df_metrics, aes(x = horizon, y = RMSE)) +
          geom_line(color = "red") + geom_point() +
          theme_minimal() +
          ggtitle("City:", city_name)
  )
}
```

The RMSE generally decreases from short horizons to intermediate horizons (around h=3–5), where the lowest errors are observed. This indicates that the model achieves its best performance in the short-to-medium term.

For longer horizons, particularly at h=7, the RMSE increases significantly, reflecting the expected accumulation of uncertainty in multi-step forecasting. Some variability is observed at lower horizons, suggesting sensitivity to short-term fluctuations.

Overall, the model demonstrates stable and reliable performance in intermediate horizons, with predictable degradation as the forecast horizon increases.

```{r, fig.width=18, fig.height=8}
# Feature Importance
for (city_name in names(out)) {
  top_features <- head(out[[city_name]]$importance_summary, 10)
  
  print(ggplot(top_features, aes(x = reorder(Feature, Gain_mean), y = Gain_mean)) +
    geom_col() +
    coord_flip() +
    ggtitle("City:", city_name) +
    theme_minimal()
  )
}
```


