---
title: "Project"
output: html_document
date: "2026-03-09"
---

# Libraries
```{r}
library(rminer)
library(aTSA)
library(dplyr)
library(timeDate)
```

# Data
```{r}
baltimore <- read.csv("data/baltimore.csv", sep = ",")
lancaster <- read.csv("data/lancaster.csv", sep = ",")
philadelphia <- read.csv("data/philadelphia.csv", sep= ",")
richmond <- read.csv("data/richmond.csv", sep = ",")
```

# EDA & Pre-Processing

```{r}
cat("--- Baltimore ---\n\n")
str(baltimore)

cat("\n--- Lancaster ---\n\n")
str(lancaster)

cat("\n--- Philadelphia ---\n\n")
str(philadelphia)

cat("\n --- Richmond ---\n\n")
str(richmond)
```

```{r}
# Convert Date
baltimore$Date <- as.Date(baltimore$Date)
lancaster$Date <- as.Date(lancaster$Date)
philadelphia$Date <- as.Date(philadelphia$Date)
richmond$Date <- as.Date(richmond$Date)
```

```{r}
cat("--- Baltimore ---\n\n")
str(baltimore)

cat("\n--- Lancaster ---\n\n")
str(lancaster)

cat("\n--- Philadelphia ---\n\n")
str(philadelphia)

cat("\n --- Richmond ---\n\n")
str(richmond)
```


```{r}
# Missing Values
cat("Number of Missing Values for Baltimore Dataset: ", sum(is.na(baltimore)), "\n")
cat("Number of Missing Values for Lancaster Dataset: ", sum(is.na(lancaster)), "\n")
cat("Number of Missing Values for Philadelphia Dataset: ", sum(is.na(philadelphia)), "\n")
cat("Number of Missing Values for Richmond Dataset: ", sum(is.na(richmond)))
```

```{r}
# Duplicated Values
cat("Number of Duplicated Values for Baltimore Dataset: ", sum(duplicated(baltimore)), "\n")
cat("Number of Duplicated Values for Lancaster Dataset: ", sum(duplicated(lancaster)), "\n")
cat("Number of Duplicated Values for Philadelphia Dataset: ", sum(duplicated(philadelphia)), "\n")
cat("Number of Duplicated Values for Richmond Dataset: ", sum(duplicated(richmond)))
```

```{r}
cat("--- Baltimore ---\n")
print(subset(baltimore, Num_Customers == 0))

cat("\n--- Lancaster ---\n")
print(subset(lancaster, Num_Customers == 0))

cat("\n--- Philadelphia ---\n")
print(subset(philadelphia, Num_Customers == 0))

cat("\n--- Richmond ---\n")
print(subset(richmond, Num_Customers == 0))
```

```{r}
cat("--- Baltimore ---\n")
print(subset(baltimore, Num_Customers == 0 & Sales > 0))

cat("\n--- Lancaster ---\n")
print(subset(lancaster, Num_Customers == 0 & Sales > 0))

cat("\n--- Philadelphia ---\n")
print(subset(philadelphia, Num_Customers == 0 & Sales > 0))

cat("\n--- Richmond ---\n")
print(subset(richmond, Num_Customers == 0 & Sales > 0))
```

From the previous analysis, we identified days where the number of customers is equal to zero. In some of these cases, namely on December 25th, it is clear that the stores were closed, as there were no employees, no customers, and no sales recorded. These observations are consistent with expected business behavior and should be retained in the dataset. It is important for the forecasting models to learn that on specific dates (e.g., Christmas), the stores are closed, leading to zero customers and zero sales.

However, we also identified cases where the number of customers is zero while sales are strictly positive. In particular, this occurs on specific dates where there are employees working and sales recorded, but no customers. This situation is logically inconsistent, since sales necessarily imply the presence of customers. Even in the case of online purchases, each transaction corresponds to a customer, and therefore the number of customers cannot be zero when sales are observed.

Since removing observations in time series can break the temporal structure and introduce bias into the modeling process, these cases are not removed. Instead, a correction is applied based on business logic: whenever the number of customers is zero, the corresponding sales values are set to zero. This ensures logical consistency in the data while preserving the temporal continuity of the series.

Additionally, missing values in variables such as *Pct_On_Sale* are handled through context-aware imputation. In cases where no customers are observed, missing promotional activity is assumed to be zero, as promotions are not meaningful in the absence of customers.

This approach preserves all observations, corrects inconsistencies, and maintains the integrity of the time series, ensuring that the dataset remains suitable for reliable forecasting.

```{r}
baltimore$Sales[baltimore$Num_Customers == 0] <- 0
lancaster$Sales[lancaster$Num_Customers == 0] <- 0
philadelphia$Sales[philadelphia$Num_Customers == 0] <- 0
richmond$Sales[richmond$Num_Customers == 0] <- 0
```

```{r}
baltimore$Pct_On_Sale[is.na(baltimore$Pct_On_Sale)] <- 0
lancaster$Pct_On_Sale[is.na(lancaster$Pct_On_Sale)] <- 0
philadelphia$Pct_On_Sale[is.na(philadelphia$Pct_On_Sale)] <- 0
richmond$Pct_On_Sale[is.na(richmond$Pct_On_Sale)] <- 0
```


```{r}
cat("--- Baltimore ---\n")
print(subset(baltimore, Num_Customers == 0 & Sales > 0))

cat("\n--- Lancaster ---\n")
print(subset(lancaster, Num_Customers == 0 & Sales > 0))

cat("\n--- Philadelphia ---\n")
print(subset(philadelphia, Num_Customers == 0 & Sales > 0))

cat("\n--- Richmond ---\n")
print(subset(richmond, Num_Customers == 0 & Sales > 0))
```

```{r}
cat("Number of Missing Values for Baltimore Dataset: ", sum(is.na(baltimore)), "\n")
cat("Number of Missing Values for Lancaster Dataset: ", sum(is.na(lancaster)), "\n")
cat("Number of Missing Values for Philadelphia Dataset: ", sum(is.na(philadelphia)), "\n")
cat("Number of Missing Values for Richmond Dataset: ", sum(is.na(richmond)))
```

```{r}
# Check the evolution of the Number of Customers
plot(baltimore$Date, baltimore$Num_Customers,
     type = "l",
     main = "Baltimore - Number of Customers",
     xlab = "Date",
     ylab = "Customers")

plot(lancaster$Date, lancaster$Num_Customers,
     type = "l",
     main = "Lancaster - Number of Customers",
     xlab = "Date",
     ylab = "Customers")

plot(philadelphia$Date, philadelphia$Num_Customers,
     type = "l",
     main = "Philadelphia - Number of Customers",
     xlab = "Date",
     ylab = "Customers")

plot(richmond$Date, richmond$Num_Customers,
     type = "l",
     main = "Richmond - Number of Customers",
     xlab = "Date",
     ylab = "Customers")
```

The time series of the number of customers shows a relatively stable level over time, with most observations ranging between approximately 80 and 200 customers per day.

There is no clear long-term trend, indicating that the overall demand remains stable throughout the observed period.

A strong weekly seasonal pattern is visible, characterized by regular fluctuations that likely correspond to differences between weekdays and weekends.

Additionally, the series presents occasional sharp spikes, where the number of customers increases significantly above the typical range. These peaks may be associated with promotional activities or special events.

Overall, the data exhibits a combination of stable baseline demand, clear weekly seasonality, and sporadic high-demand events.

```{r}
# Check the Stationary
cat("--- Baltimore ---\n")
adf.test(baltimore$Num_Customers)
cat("\n--- Lancaster ---\n")
adf.test(lancaster$Num_Customers)
cat("\n--- Philadelphia ---\n")
adf.test(philadelphia$Num_Customers)
cat("\n--- Richmond ---\n")
adf.test(richmond$Num_Customers)
```

The `adf.test()` was applied to the target variable, *Num_Customers*, to access stationarity. The results show a p-value <= 0.01, allowing us to reject the null hypothesis of non-stationarity. Therefore, the series can be considered **stationarity**.

Given this result, no differencing is required, and the analysis can be proceed using the original serires.

```{r}
# Check Autocorrelation
acf(baltimore$Num_Customers)
acf(lancaster$Num_Customers)
acf(philadelphia$Num_Customers)
acf(richmond$Num_Customers)
```

```{r}
# Check Autocorrelation with differencing
acf(diff(baltimore$Num_Customers))
acf(diff(lancaster$Num_Customers))
acf(diff(philadelphia$Num_Customers))
acf(diff(richmond$Num_Customers))
```

The ACF analysis of the original series reveals a strong dependence between consecutive observations, particularly at lag 1, along with clear seasonal patterns. Significant spikes at lag 7 and its multiples (14, 21, 28) indicate a consistent weekly seasonality across all stores.

However, when analyzing the differenced series, most autocorrelations fall within the confidence bounds, suggesting that a large part of this dependence was driven by underlying temporal structures such as trend and seasonality. Only a few lags (notably lag 1 and, to a lesser extent, lag 7) remain slightly significant, indicating limited true temporal dependence.

Overall, while the original series suggests strong autocorrelation and seasonality, the differenced analysis shows that the effective dependency structure is much weaker, reinforcing that only a small number of lags are relevant for modeling.

```{r}
# Partial Autocorrelation
pacf(baltimore$Num_Customers)
pacf(lancaster$Num_Customers)
pacf(philadelphia$Num_Customers)
pacf(richmond$Num_Customers)
```

```{r}
# Partial Autocorrelation with differencing
pacf(diff(baltimore$Num_Customers))
pacf(diff(lancaster$Num_Customers))
pacf(diff(philadelphia$Num_Customers))
pacf(diff(richmond$Num_Customers))
```

The PACF analysis of the original series shows several significant partial autocorrelations across multiple lags, suggesting a complex dependency structure. This pattern is likely influenced by trend and seasonality effects, which inflate the number of seemingly relevant lags.

After applying differencing, the PACF structure becomes notably simpler. Most lags fall within the confidence bounds, with only the first few lags (particularly lag 1) showing some residual significance. This indicates that the true direct dependence is limited and that much of the structure observed in the original series was driven by non-stationary components.

Overall, while the original PACF suggests a higher-order autoregressive behavior, the differenced series reveals that only a small number of lags are truly relevant, supporting a more parsimonious model specification.

```{r}
# Check if Promotions affect customers who are late
ccf(baltimore$Pct_On_Sale, baltimore$Num_Customers)
ccf(lancaster$Pct_On_Sale, lancaster$Num_Customers)
ccf(philadelphia$Pct_On_Sale, philadelphia$Num_Customers)
ccf(richmond$Pct_On_Sale, richmond$Num_Customers)
```
While negative lags could theoretically capture anticipation effects (e.g., customers delaying purchases in expectation of future promotions), the current dataset does not include information about events or promotions known in advance. Therefore, negative lags are not considered meaningful for this analysis.

```{r}
ccf_positive <- function(x, y, lag.max = 28, main = "CCF (Positive Lags Only)") {
  if (anyNA(x) || anyNA(y)) {
    stop("x or y has missing values. Please handle missing values before using `ccf_positivie()`.")
  }
  
  if (!is.numeric(x) || !is.numeric(y)) {
    stop("x and y must be numeric values.")
  }
  
  if (length(x) != length(y)) {
    stop("x and y must have the same length.")
  }
  
  # Calculate CCf without plot
  ccf_obj <- ccf(x, y, lag.max = lag.max, plot = FALSE)
  
  # Get the values
  lags <- as.vector(ccf_obj$lag)
  acf_vals <- as.vector(ccf_obj$acf)
  
  # Get the positive lags
  pos_index <- lags >= 0
  
  lags_pos <- lags[pos_index]
  acf_pos <- acf_vals[pos_index]
  
  # Calculate confidence interval
  n <- length(x)
  conf <- 2 / sqrt(n)
  
  plot(lags_pos, acf_pos, type = "h",
       xlab = "Lag", ylab = "ACF",
       main = main)
  
  abline(h = 0, col = "black")
  abline(h = c(-conf, conf), col = "blue", lty = 2)
}
```

This function was created to check only the positive values of LAG.

```{r}
# Check if Promotions affect customers who are late
ccf_positive(baltimore$Pct_On_Sale, baltimore$Num_Customers)
ccf_positive(lancaster$Pct_On_Sale, lancaster$Num_Customers)
ccf_positive(philadelphia$Pct_On_Sale, philadelphia$Num_Customers)
ccf_positive(richmond$Pct_On_Sale, richmond$Num_Customers)
```


```{r}
# Check if Promotions affect customers who are late. With Differencing
ccf_positive(diff(baltimore$Pct_On_Sale), diff(baltimore$Num_Customers))
ccf_positive(diff(lancaster$Pct_On_Sale), diff(lancaster$Num_Customers))
ccf_positive(diff(philadelphia$Pct_On_Sale), diff(philadelphia$Num_Customers))
ccf_positive(diff(richmond$Pct_On_Sale), diff(richmond$Num_Customers))
```

The **CCF (Cross-Correlation Function)** between *Pct_On_Sale* and *Num_Customers* initially shows several positive correlations across multiple lags, with a clear peak at Lag0. This suggests that promotions are associated with an increase in customers both contemporaneously and across different time shifts.

However, the presence of many significant lags indicates that this pattern may be driven by shared trend or seasonality rather than a true causal relationship. In other words, both variables may be evolving similarly over time, leading to spurious correlations that can mislead interpretation.

After applying `diff()`, the analysis focuses on changes between consecutive observations, effectively removing trend and reducing the impact of seasonality. As a result, the CCF becomes much cleaner, with most of the previously significant lags disappearing.

The peak at Lag0 remains clearly significant, confirming an immediate effect of promotions on customer numbers. Additionally, only a small number of short-term lags (e.g., Lag1 or Lag2) may retain some relevance, suggesting a limited delayed impact.

Overall, while the original CCF highlights general co-movement between the variables, the differenced CCF isolates the true dynamic relationship. This makes the version with `diff()` more reliable for identifying meaningful effects and selecting appropriate lags for modeling.

```{r}
# Check if the Events have an immediate or delayed effect

# Create a new numerical variable from TouristEvent (just for analysis purposes)
baltimore$TouristEvent_num <- ifelse(baltimore$TouristEvent == 'Yes', 1, 0)
lancaster$TouristEvent_num <- ifelse(lancaster$TouristEvent == 'Yes', 1, 0)
philadelphia$TouristEvent_num <- ifelse(philadelphia$TouristEvent == 'Yes', 1, 0)
richmond$TouristEvent_num <- ifelse(richmond$TouristEvent == 'Yes', 1, 0)

# CCF
ccf_positive(baltimore$TouristEvent_num, baltimore$Num_Customers)
ccf_positive(lancaster$TouristEvent_num, lancaster$Num_Customers)
ccf_positive(philadelphia$TouristEvent_num, philadelphia$Num_Customers)
ccf_positive(richmond$TouristEvent_num, richmond$Num_Customers)
```



```{r}
# Check if the Events have an immediate or delayed effect. With differencing
ccf_positive(diff(baltimore$TouristEvent_num), diff(baltimore$Num_Customers))
ccf_positive(diff(lancaster$TouristEvent_num), diff(lancaster$Num_Customers))
ccf_positive(diff(philadelphia$TouristEvent_num), diff(philadelphia$Num_Customers))
ccf_positive(diff(richmond$TouristEvent_num), diff(richmond$Num_Customers))
```

The **CCF (Cross-Correlation Function)** between *TouristEvent_num* and *Num_Customers* shows a strong and consistent peak at Lag0 across all locations, indicating a clear contemporaneous relationship: the presence of events is directly associated with an increase in the number of customers on the same day.

In the non-differenced series, this effect appears very pronounced, with high correlations at Lag0 and some neighboring lags. However, the presence of additional significant correlations around nearby lags suggests that part of this relationship may be influenced by shared temporal structures, such as trend or recurring patterns, potentially inflating the apparent impact.

After applying `diff()`, the overall structure becomes more refined. The dominant peak at Lag0 remains strong and statistically significant, confirming that the relationship is not spurious and reflects a genuine immediate effect of events on customer demand. At the same time, most of the surrounding correlations are reduced or disappear, indicating that they were largely driven by underlying temporal dependencies rather than true lagged effects.

Importantly, no consistent or strong delayed effects (Lag > 0) are observed after differencing, suggesting that events primarily influence customer numbers instantaneously, with little to no persistence over subsequent periods.

This pattern is consistent across all stores, reinforcing the conclusion that events act as an immediate demand driver, and that incorporating lagged versions of this variable is not necessary for modeling purposes.

```{r}
# Create LAG, 1 and 7 (LAG7 we going to create even though the results of ACF and PACF told us that it is not that good, so the creation is only for testing purposes, and if the predictions got better this variable stays)
baltimore$lag_1 <- lag(baltimore$Num_Customers, 1)
baltimore$lag_2 <- lag(baltimore$Num_Customers, 2)
baltimore$lag_7 <- lag(baltimore$Num_Customers, 7)
baltimore <- na.omit(baltimore)

lancaster$lag_1 <- lag(lancaster$Num_Customers, 1)
lancaster$lag_2 <- lag(lancaster$Num_Customers, 2)
lancaster$lag_7 <- lag(lancaster$Num_Customers, 7)
lancaster <- na.omit(lancaster)

philadelphia$lag_1 <- lag(philadelphia$Num_Customers, 1)
philadelphia$lag_2 <- lag(philadelphia$Num_Customers, 2)
philadelphia$lag_7 <- lag(philadelphia$Num_Customers, 7)
philadelphia <- na.omit(philadelphia)

richmond$lag_1 <- lag(richmond$Num_Customers, 1)
richmond$lag_2 <- lag(richmond$Num_Customers, 2)
richmond$lag_7 <- lag(richmond$Num_Customers, 7)
richmond <- na.omit(richmond)
```

The descriptive statistics reveal clear differences in scale between the stores. Philadelphia and Lancaster exhibit higher average numbers of customers and sales, while Richmond consistently shows lower values, indicating smaller demand.

Across all stores, the maximum values are substantially higher than the median, suggesting the presence of significant peaks, likely associated with promotional events or external factors.

The distributions of both customers and sales appear to be positively skewed, as the mean tends to exceed the median. Additionally, the presence of zero values confirms the existence of closed days or anomalous observations previously identified.

Overall, while the stores share similar behavioral patterns, they differ considerably in magnitude, reinforcing the need to account for store-specific effects in the modeling phase.

```{r}
# Check if the Tourist Event has impact in the Number of Customers
boxplot(Num_Customers ~ TouristEvent, data = baltimore)
title("Baltimore")
boxplot(Num_Customers ~ TouristEvent, data = lancaster)
title("Lancaster")
boxplot(Num_Customers ~ TouristEvent, data = philadelphia)
title("Philadelphia")
boxplot(Num_Customers ~ TouristEvent, data = richmond)
title("Richmond")
```

This **Boxplot** shows a substantial increase in the number of customers during tourits events across all stores. The distributions for event days are significantly shifted upward compared to non-event days, with the minimal overlap between the two groups.

This indicates that tourist events are a major driver of customer demand and should be considered a key explanatory variable in the forecasting model.

```{r}
# Check if the Tourist Event has impact in Sales
boxplot(Sales ~ TouristEvent, data = baltimore)
title("Baltimore")
boxplot(Sales ~ TouristEvent, data = lancaster)
title("Lancaster")
boxplot(Sales ~ TouristEvent, data = philadelphia)
title("Philadelphia")
boxplot(Sales ~ TouristEvent, data = richmond)
title("Richmond")
```

```{r}
# Check if the TouristEvent increase quantity or average ticket

baltimore$Sales_per_Customer <- baltimore$Sales / baltimore$Num_Customers
lancaster$Sales_per_Customer <- lancaster$Sales / lancaster$Num_Customers
philadelphia$Sales_per_Customer <- philadelphia$Sales / philadelphia$Num_Customers
richmond$Sales_per_Customer <- richmond$Sales / richmond$Num_Customers

boxplot(Sales_per_Customer ~ TouristEvent, data = baltimore)
title("Baltimore")
boxplot(Sales_per_Customer ~ TouristEvent, data = lancaster)
title("Lancaster")
boxplot(Sales_per_Customer ~ TouristEvent, data = philadelphia)
title("Philadelphia")
boxplot(Sales_per_Customer ~ TouristEvent, data = richmond)
title("Richmond")
```

In this code cell, a new feature *Sales_per_Customer* was created to further investigate the impact of *TouristEvent* on customer behavior. Although this feature provides valuable insights, revealing that tourists events increase the number of customers while decreasing the average sales per customer, it cannot be used for modeling purposes. This is because it depends on the target variable (Num_Customer) and on information not available at prediction time, which would introduce data leakage and lead to unrealistic model performance.

Therefore, this feature is retained only for exploratory analysis and removed from the modeling dataset. Additionally, *TouristEvent* is treated as a categorical Variable, as it plays a key role in explaining variability in the data, including peak and apparent outliers.

```{r}
baltimore$TouristEvent <- as.factor(baltimore$TouristEvent)
lancaster$TouristEvent <- as.factor(lancaster$TouristEvent)
philadelphia$TouristEvent <- as.factor(philadelphia$TouristEvent)
richmond$TouristEvent <- as.factor(richmond$TouristEvent)
```


```{r}
create_holiday_feature <- function(df) {
  # This functions is only used to create holiday feature for US Calendar
  df$Date <- as.Date(df$Date)

  start_year <- as.numeric(format(min(df$Date), "%Y"))
  end_year <- as.numeric(format(max(df$Date), "%Y"))

  years <- start_year:end_year

  us_holidays <- as.Date(holidayNYSE(years))

  thanksgiving <- as.Date(USThanksgivingDay(years))
  black_friday <- as.Date(thanksgiving + 1)
  christmas <- as.Date(USChristmasDay(years))

  all_holidays <- unique(c(us_holidays, thanksgiving, black_friday, christmas))

  df$IsHoliday <- ifelse(df$Date %in% all_holidays, 1, 0)
  df$IsBlackFriday <- ifelse(df$Date %in% black_friday, 1, 0)
  df$IsChristmas <- ifelse(df$Date %in% christmas, 1, 0)

  return(df)
}
```

```{r}
baltimore <- create_holiday_feature(baltimore)
lancaster <- create_holiday_feature(lancaster)
philadelphia <- create_holiday_feature(philadelphia)
richmond <- create_holiday_feature(richmond)
```

Feature engineering plays a crucial role in improving the model’s ability to capture underlying patterns in time series data. In this context, additional calendar-based features were introduced to incorporate external information that is not directly observable from the original variables.

Specifically, holiday-related variables were created, including general holidays (`IsHoliday`), Christmas (`IsChristmas`), and Black Friday (`IsBlackFriday`). These features allow the model to distinguish between different types of days that have distinct impacts on customer behavior. For example, Christmas is associated with store closures and zero activity, while Black Friday typically leads to unusually high demand.

By explicitly providing this contextual information, the model can better interpret extreme values, reduce ambiguity in the data, and improve its predictive performance.

```{r}
cat("--- Baltimore ---\n\n")
summary(baltimore)

cat("\n--- Lancaster ---\n\n")
summary(lancaster)

cat("\n--- Philadelphia ---\n\n")
summary(philadelphia)

cat("\n--- Richmond ---\n\n")
summary(richmond)
```

### *Detect Outliers*

Outlier analysis is an essential part of Exploratory Data Analysis. While some observations appear as statistical outliers, many of them are driven by real-world events and therefore represent meaningful behavior rather than anomalies. As such, these values are retained to preserve the integrity of the time series.

In time series analysis, observations are not removed based solely on their extremity, as doing so would disrupt the temporal structure of the data. Instead, each case is evaluated based on business logic. Extreme values associated with real events (e.g., promotions or holidays) are preserved, while only logically inconsistent observations are corrected or imputed to maintain data integrity without breaking the time sequence.

```{r}
boxplot(baltimore$Sales, ylab = "Sales", xlab = "Count", main = "Baltimore", horizontal = TRUE)
boxplot(lancaster$Sales, ylab = "Sales", xlab = "Count", main = "Baltimore", horizontal = TRUE)
boxplot(philadelphia$Sales, ylab = "Sales", xlab = "Count", main = "Baltimore", horizontal = TRUE)
boxplot(richmond$Sales, ylab = "Sales", xlab = "Count", main = "Baltimore", horizontal = TRUE)
```

```{r}
# Get indexes where boxplot tells which observations are outliers
get_outliers <- function(df, var) {
  if (anyNA(df[[var]])) {
    stop("This Feature has Missing Values. Please handle first the Missing Values.")
  }
  x <- df[[var]]
  
  # Calculate Boxplot Outliers
  Q1 <- quantile(x, 0.25)
  Q3 <- quantile(x, 0.75)
  IQR <- Q3-Q1
  
  upper_bound = Q3 + 1.5 * IQR
  lower_bound = Q1 - 1.5 * IQR
  
  boxplot_outliers <- which(x < lower_bound | x > upper_bound)
    
  # Calculate Hampel Identifier Outliers
  median_x <- median(x)
  mad_x <- mad(x, constant = 1)
  
  # Threshold ~ 3
  hampel_threshold <- 3 * mad_x
  hampel_outliers <- which(abs(x - median_x) > hampel_threshold)
  
  # Intersection between metrics
  intersection_outliers <- intersect(boxplot_outliers, hampel_outliers)
  
  return(list(
    iqrOutlier = boxplot_outliers,
    hampelOutlier = hampel_outliers,
    intersectionOutlier = intersection_outliers
  ))
}
```

```{r}
# Check outliers
outlier <- get_outliers(baltimore, "Sales")
```

```{r}
cat("--- IQR Outliers ---\n")
cat("Outliers:", outlier$iqrOutlier, "\n")
cat("Number of Outliers:", length(outlier$iqrOutlier), "\n")

cat("\n--- Hampel Outliers ---\n")
cat("Outliers:", outlier$hampelOutlier, "\n")
cat("Number of Outliers:", length(outlier$hampelOutlier), "\n")

cat("\n--- Intersection ---\n")
cat("Outliers:", outlier$intersectionOutlier, "\n")
cat("Number of Outliers:", length(outlier$intersectionOutlier))

if (identical(outlier$iqrOutlier, outlier$intersectionOutlier)) {
  cat("\nIntersection is exactly equal to IQR!")
} else {
  cat("\nIntersection is a subset of both!")
}
```


